---
title             : "Understanding how U.S. schools and school districts used Facebook to communicate before and during the COVID-19 pandemic"
shorttitle        : "Title"

author: 
  - name          : "First Author"
    affiliation   : "1"
    corresponding : yes    # Define only one corresponding author
    address       : "Postal address"
    email         : "my@email.com"
  - name          : "Ernst-August Doelle"
    affiliation   : "1,2"

affiliation:
  - id            : "1"
    institution   : "Wilhelm-Wundt-University"
  - id            : "2"
    institution   : "Konstanz Business School"

authornote: |
  Add complete departmental affiliations for each author here. Each new line herein must be indented, like this line.

  Enter author note here.

abstract: |
  One or two sentences providing a **basic introduction** to the field,  comprehensible to a scientist in any discipline.
  
  Two to three sentences of **more detailed background**, comprehensible  to scientists in related disciplines.
  
  One sentence clearly stating the **general problem** being addressed by  this particular study.
  
  One sentence summarizing the main result (with the words "**here we show**" or their equivalent).
  
  Two or three sentences explaining what the **main result** reveals in direct comparison to what was thought to be the case previously, or how the  main result adds to previous knowledge.
  
  One or two sentences to put the results into a more **general context**.
  
  Two or three sentences to provide a **broader perspective**, readily comprehensible to a scientist in any discipline.
  
  <!-- https://tinyurl.com/ybremelq -->
  
keywords          : "keywords"
wordcount         : "X"

bibliography      : ["r-references.bib"]

floatsintext      : no
figurelist        : no
tablelist         : no
footnotelist      : no
linenumbers       : yes
mask              : no
draft             : no

documentclass     : "apa6"
classoption       : "man"
output            : papaja::apa6_pdf
---

```{r setup, include = FALSE}
library("papaja")
r_refs("r-references.bib")

```

```{r analysis-preferences}
# Seed for random number generation
set.seed(42)
knitr::opts_chunk$set(cache.extra = knitr::rand_seed)
```

Our purpose is to understand how U.S. schools and school districts used Facebook to communicate before and during the COVID-19 pandemic. We do so using a data source of XXX posts from schools and districts from XXX accounts on Facebook in the 2019 and 2020 years. 

Our research questions are:

1. How did the COVID-19 pandemic impact the public-facing communication strategies of schools and school districts? 

This is important to answer because this can help us:
- to understand what districts said to THE PUBLIC (who are the recipients of these?)
- be able to speak to emergent best practices for districts’ social media use. 
- this can tell us about how they're responding to COVID (so using this data as a proxy for their response)

Specifically, we examine the topics of posts

- Look at how these are interacted with differently
- Look at frequency of the topics of posts

*2. What variation do we see in their use?*

# Methods
We report how we determined our sample size, all data exclusions (if any), all manipulations, and all measures in the study. <!-- 21-word solution (Simmons, Nelson & Simonsohn, 2012; retrieved from http://ssrn.com/abstract=2160588) -->

The number of topics (k = X) was determined based on several model fit diagnostics, including held-out likelihood, residual check, exclusivity, and semantic coherence. Higher held-out likelihood suggests higher predictive power. For a document subset, the algorithm removes half of the set, and calculates the probability of detecting the held out words. We selected topic numbers with lower residual checks, which suggest that fewer topics would be necessary to explain data variance (Roberts et al., 2014). Finally, the choice of topic number achieved a balance between high semantic coherence and high exclusivity. Semantic coherence indicates the consistency of the predicted topics, such that words that appear frequently in a topic are more likely to co-occur within documents. Exclusivity describes the uniqueness of a topic's high-frequency words, such that these words do not overlap with the most frequent words of other topics.

## Participants

## Material

## Procedure

## Data analysis
We used `r cite_r("r-references.bib")` for all our analyses.


# Results

# Discussion


\newpage

# References

\begingroup
\setlength{\parindent}{-0.5in}
\setlength{\leftskip}{0.5in}

<div id="refs" custom-style="Bibliography"></div>
\endgroup
