---
title: "Initial Exploration of FB Posts Database"
date: 2020-10-21
output:
  html_document:
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, dpi = 250, warning = FALSE)
knitr::clean_cache()
```

# Loading, setting up

```{r, load-packages-and-connect-to-db}
library(tidyverse) # analysis and processing
library(DBI) # database interface
library(scales) # plots
library(quanteda) # text analysis

cn <- dbConnect(RSQLite::SQLite(), dbname = here::here("db", "k12-institutions-fb-posts.sqlite"))
```

## Total number of posts

```{r}
total_unique_posts <- tbl(cn, "posts") %>%
  tally()
```

URL is a unique identifier

```{r, inspect-table}
tbl(cn, "posts") %>% 
  count(url) %>%
  tally()
```

# Types of post

```{r}
tbl(cn, "posts") %>% 
  count(type) %>% 
  arrange(desc(n)) %>% 
  knitr::kable()
```

# Accounts

## Total number of accounts

```{r, n-accounts}
tbl(cn, "posts") %>% 
  count(facebook_id) %>% 
  tally()
```

## Distribution of the number of posts by accounts

```{r, users-dist}
tbl(cn, "posts") %>% count(facebook_id) %>% 
  ggplot(aes(x = n)) + 
  geom_histogram(bins = 1000)
```

# Time series

```{r, plot-between-years}
by_day <- tbl(cn, "posts") %>% 
  count(created_rounded_to_day) %>% 
  collect() %>% 
  mutate(day = as.POSIXct(created_rounded_to_day, origin = "1970-01-01"))
```

## Month of year - within the year

```{r, plot-within-year}
by_day %>% 
  mutate(yday = lubridate::yday(day),
         month = lubridate::month(day),
         year = lubridate::year(day)) %>% 
  filter(year >= 2009) %>% 
  mutate(year = as.factor(year)) %>% 
  group_by(year, month) %>% 
  summarize(sum_n = sum(n)) %>% 
  ggplot(aes(x = month, y = sum_n, color = year, group = year)) +
  geom_point() +
  geom_line() +
  hrbrthemes::theme_ipsum(base_size = 13) +
  scale_y_continuous(label = comma) +
  scale_x_continuous(labels = 1:12, breaks = 1:12) +
  scale_color_discrete("",guide = guide_legend(reverse = TRUE)) +
  ylab("Posts") +
  xlab("Month")

ggsave('posts-per-month.png', width = 8, height = 6)
```

## Month of year - between years

```{r, plot-month}
by_day %>% 
  ggplot(aes(x = day, y = n)) + 
  geom_point(alpha = .25) +
  geom_line() +
  theme_bw() +
  scale_y_continuous(label = comma) +
  geom_smooth(se = FALSE, color = "blue")
```

## Day of week

```{r, plot-day}
tbl(cn, "posts") %>%  
  count(day_of_week) %>% 
  collect() %>% 
  mutate(day_of_week_string = c("Sunday", "Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday")) %>% 
  mutate(day_of_week_string = fct_inorder(day_of_week_string)) %>% 
  ggplot(aes(x = day_of_week_string, y = n)) +
  geom_col() +
  theme_bw() + 
  scale_y_continuous(label = comma)
```

## Hour

```{r, plot-hour}
tbl(cn, "posts") %>%  
  count(hour) %>% 
  collect() %>% 
  mutate(hour = lubridate::make_datetime(hour = hour)) %>% 
  ggplot(aes(x = hour, y = n)) +
  geom_col() +
  hrbrthemes::theme_ipsum() +
  scale_x_datetime(date_labels = "%H:%M") +
  scale_y_continuous(label = comma)
```

# Interactions

## Total interactions per post (overall)

```{r, total-interactions-per-post}
tbl(cn, "posts") %>% 
  select(likes:care) %>% 
  summarize_all(funs(mean, sd)) %>% 
  collect() %>% 
  gather(key, val) %>% 
  summarize(mean_total_interactions_per_post = sum(val),
            sd_total_interactions_per_post = sd(val))
```

## Mean total interactions per post (overall)

```{r, mean-total-interactions}
tbl(cn, "posts") %>% 
  select(likes:care, year) %>% 
  summarize_all(funs(mean, sd)) %>% 
  collect() %>% 
  select(-year_mean, -year_sd) %>% 
  gather(key, val) %>% 
  filter(!str_detect(key, "sd")) %>%
  summarize(mean_total_interactions_per_post = sum(val),
            sd_total_interactions_per_post = sd(val))
```

## Another way of presenting interactions

probably preferable to and redundant with the others for the purposes of exploring the data, at least the others that are not by time

```{r, descriptives-of-interactions-prep}
descriptives_of_interactions <- tbl(cn, "posts") %>% 
  select(likes:care) %>% 
  summarize_all(funs(sum, mean, sd)) %>% 
  collect()
```

```{r, descriptives-of-interactions-plot}
table_of_descriptives <- descriptives_of_interactions %>% 
  gather(key, val) %>% 
  separate(key, into = c("var", "stat")) %>% 
  spread(stat, val) %>% 
  arrange(desc(sum))

table_of_descriptives %>% 
  knitr::kable()
```

```{r, descriptives-sum}
table_of_descriptives %>% 
  summarize_if(is.numeric, sum)
```

## Mean interactions by type per post (overall)

```{r, mean-interactions-by-type-of-post-plot}
interactions <- tbl(cn, "posts") %>% 
  select(likes:care) %>% 
  summarize_all(funs(mean, sd)) %>% 
  collect()

interactions %>% 
  select(contains("mean")) %>% 
  gather(key, val) %>% 
  ggplot(aes(x = reorder(key,val), y = val)) + 
  geom_col() + 
  coord_flip() +
  geom_smooth(se = FALSE) + 
  hrbrthemes::theme_ipsum()
```

## Mean interactions by type per post by Month (2010-September, 2020)

```{r, mean-interactions-by-type-by-month}
interactions_by_month <- tbl(cn, "posts") %>% 
  group_by(year, month) %>% 
  select(likes:care) %>% 
  summarize_all(funs(mean, sd)) %>% 
  collect()

interactions_by_month %>% 
  select(contains("mean"), year, month) %>%
  mutate(date = lubridate::make_date(year = year, month = month)) %>% 
  ungroup() %>% 
  select(-year, -month) %>% 
  gather(key, val, -date) %>% 
  filter(date >= lubridate::ymd("2010-01-01")) %>% 
  filter(key != "care_mean", key != "angry_mean", key != "wow_mean") %>% 
  ggplot(aes(x = date, y= val, group = key, color = key)) +
  geom_point(alpha = .2) +
  geom_line() +
  geom_smooth(se = FALSE) + 
  hrbrthemes::theme_ipsum()
```

## Mean interactions by type by month (2018-September, 2020)

```{r, mean-interactions-by-type-by-month-focused}
interactions_by_month %>% 
  select(contains("mean"), year, month) %>%
  mutate(date = lubridate::make_date(year = year, month = month)) %>% 
  ungroup() %>% 
  select(-year, -month) %>% 
  gather(key, val, -date) %>% 
  filter(date >= lubridate::ymd("2018-01-01")) %>% 
  filter(key != "care_mean", key != "angry_mean", key != "wow_mean") %>% 
  ggplot(aes(x = date, y= val, group = key, color = key)) +
  geom_point(alpha = .2) +
  geom_line() +
  geom_smooth(se = FALSE) + 
  hrbrthemes::theme_ipsum()
```