---
title: "Joining NCES IDs"
date: 2020-11-12
output:
  html_document:
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, dpi = 250, warning = FALSE)
knitr::clean_cache()
```

# Loading, setting up

```{r, load-packages-and-connect-to-db}
library(tidyverse) # analysis and processing
library(DBI) # database interface
library(scales) # plots
library(quanteda) # text analysis
library(skimr) # for descriptive stats

cn <- dbConnect(RSQLite::SQLite(), dbname = here::here("db", "k12-institutions-fb-posts.sqlite"))
```

# Let's join!

## getting data

```{r}
posts_100k <- tbl(cn, "posts") %>% 
  select(page_name, url, year, message) %>% 
  mutate(url = tolower(url)) %>% 
  filter(year == 2020) %>% 
  collect()

posts_100k <- posts_100k %>% 
  separate(url, into = c("url", "post"), sep = "/posts")

posts_100k <- posts_100k %>% 
  distinct(url, post, .keep_all = TRUE)
```

```{r}
all_institutional_facebook_urls <- read_csv("data-raw/all-institutional-facebook-urls.csv",
                                            col_types = cols(nces_id = col_character()))

all_institutional_facebook_urls <- all_institutional_facebook_urls %>% 
  mutate(parsed_path = ifelse(str_sub(parsed_path, start = -1) == "/",
                              str_sub(parsed_path, end = -2),
                              parsed_path)) %>% 
  mutate(url = str_c("https://www.facebook.com/", parsed_path)) %>% 
  select(-parsed_path)
```

```{r}
all_distinct_urls <- all_institutional_facebook_urls %>% 
  distinct(url, .keep_all = T)

data_with_nces_id_joined <- left_join(posts_100k, all_distinct_urls, by = "url")
```

## initial join

```{r}
data_with_nces_id_joined_not_matched <- data_with_nces_id_joined %>% 
  filter(is.na(nces_id))

data_with_nces_id_joined_match <- data_with_nces_id_joined %>% 
  filter(!is.na(nces_id))

str_c(round(nrow(data_with_nces_id_joined_match) / nrow(posts_100k) * 100, 2), "% joined")
```

## joining on other parts of the URL

```{r}
# 4th
fourth_spot <- all_distinct_urls %>% 
  mutate(split_url = str_split(url, "/")) %>% 
  pull(split_url) %>% 
  map(possibly(~.[[4]], NA))

all_institutional_facebook_urls_fourth <- all_distinct_urls %>% 
  mutate(url = str_c("https://www.facebook.com/", unlist(fourth_spot))) %>% 
  filter(!is.na(url))

joined_fourth <- data_with_nces_id_joined_not_matched %>% 
  select(-nces_id) %>% 
  left_join(all_institutional_facebook_urls_fourth)  %>% 
  filter(!is.na(nces_id))

# 5th
fifth_spot <- all_distinct_urls %>% 
  mutate(split_url = str_split(url, "/")) %>% 
  pull(split_url) %>% 
  map(possibly(~.[[5]], NA))

all_institutional_facebook_urls_fifth <- all_distinct_urls %>% 
  mutate(url = str_c("https://www.facebook.com/", unlist(fifth_spot))) %>% 
  filter(!is.na(url))

joined_fifth <- data_with_nces_id_joined_not_matched %>% 
  select(-nces_id) %>% 
  left_join(all_institutional_facebook_urls_fifth)  %>% 
  filter(!is.na(nces_id))

# 6th
sixth_spot <- all_distinct_urls %>% 
  mutate(split_url = str_split(url, "/")) %>% 
  pull(split_url) %>% 
  map(possibly(~.[[6]], NA))

all_institutional_facebook_urls_sixth <- all_distinct_urls %>% 
  mutate(url = str_c("https://www.facebook.com/", unlist(sixth_spot))) %>% 
  filter(!is.na(url))

joined_sixth <- data_with_nces_id_joined_not_matched %>% 
  select(-nces_id) %>% 
  left_join(all_institutional_facebook_urls_sixth) %>% 
  filter(!is.na(nces_id))

# 7th
seventh_spot <- all_distinct_urls %>% 
  mutate(split_url = str_split(url, "/")) %>% 
  pull(split_url) %>% 
  map(possibly(~.[[7]], NA))

all_institutional_facebook_urls_seventh <- all_distinct_urls %>% 
  mutate(url = str_c("https://www.facebook.com/", unlist(seventh_spot))) %>% 
  filter(!is.na(url))

joined_seventh <- data_with_nces_id_joined_not_matched %>% 
  select(-nces_id) %>% 
  left_join(all_institutional_facebook_urls_seventh) %>% 
  filter(!is.na(nces_id))

# 8th
eighth_spot <- all_distinct_urls %>% 
  mutate(split_url = str_split(url, "/")) %>% 
  pull(split_url) %>% 
  map(possibly(~.[[8]], NA))

all_institutional_facebook_urls_eighth <- all_distinct_urls %>% 
  mutate(url = str_c("https://www.facebook.com/", unlist(eighth_spot))) %>% 
  filter(!is.na(url))

joined_eighth <- data_with_nces_id_joined_not_matched %>% 
  select(-nces_id) %>% 
  left_join(all_institutional_facebook_urls_eighth)  %>% 
  filter(!is.na(nces_id))

# 9th
ninth_spot <- all_distinct_urls %>% 
  mutate(split_url = str_split(url, "/")) %>% 
  pull(split_url) %>% 
  map(possibly(~.[[9]], NA))

all_institutional_facebook_urls_ninth <- all_distinct_urls %>% 
  mutate(url = str_c("https://www.facebook.com/", unlist(ninth_spot))) %>% 
  filter(!is.na(url))

joined_ninth <- data_with_nces_id_joined_not_matched %>% 
  select(-nces_id) %>% 
  left_join(all_institutional_facebook_urls_ninth) %>% 
  filter(!is.na(nces_id))

new_joined_data <- bind_rows(joined_fourth) %>% 
  bind_rows(joined_fifth) %>% 
  bind_rows(joined_sixth) %>% 
  bind_rows(joined_seventh) %>% 
  bind_rows(joined_eighth) %>%
  bind_rows(joined_ninth) %>% 
  distinct(page_name, post, .keep_all = TRUE)
```

## removing spaces

```{r}
# fourth
all_institutional_facebook_urls_fourth_no_dash <- all_institutional_facebook_urls_fourth %>% 
  mutate(url = str_remove_all(url, "-"))

joined_fourth_no_dash <- data_with_nces_id_joined_not_matched %>% 
  select(-nces_id) %>% 
  left_join(all_institutional_facebook_urls_fourth_no_dash) %>% 
  filter(!is.na(nces_id))

# fifth
all_institutional_facebook_urls_fifth_no_dash <- all_institutional_facebook_urls_fifth %>% 
  mutate(url = str_remove_all(url, "-"))

joined_fifth_no_dash <- data_with_nces_id_joined_not_matched %>% 
  select(-nces_id) %>% 
  left_join(all_institutional_facebook_urls_fifth_no_dash) %>% 
  filter(!is.na(nces_id))

# sixth
all_institutional_facebook_urls_sixth_no_dash <- all_institutional_facebook_urls_sixth %>% 
  mutate(url = str_remove_all(url, "-"))

joined_sixth_no_dash <- data_with_nces_id_joined_not_matched %>% 
  select(-nces_id) %>% 
  left_join(all_institutional_facebook_urls_sixth_no_dash) %>% 
  filter(!is.na(nces_id))

# seventh
all_institutional_facebook_urls_seventh_no_dash <- all_institutional_facebook_urls_seventh %>% 
  mutate(url = str_remove_all(url, "-"))

joined_seventh_no_dash <- data_with_nces_id_joined_not_matched %>% 
  select(-nces_id) %>% 
  left_join(all_institutional_facebook_urls_seventh_no_dash) %>% 
  filter(!is.na(nces_id))

# eighth
all_institutional_facebook_urls_eighth_no_dash <- all_institutional_facebook_urls_eighth %>% 
  mutate(url = str_remove_all(url, "-"))

joined_eighth_no_dash <- data_with_nces_id_joined_not_matched %>% 
  select(-nces_id) %>% 
  left_join(all_institutional_facebook_urls_eighth_no_dash) %>% 
  filter(!is.na(nces_id))

new_data_without_dashes <- bind_rows(joined_fourth_no_dash) %>% 
  bind_rows(joined_fifth_no_dash) %>% 
  bind_rows(joined_sixth_no_dash) %>% 
  bind_rows(joined_seventh_no_dash) %>% 
  bind_rows(joined_eighth_no_dash) %>% 
  distinct(page_name, post, .keep_all = TRUE)
```

## joining based only on integers

```{r}
all_institutional_facebook_urls_b <- all_distinct_urls

all_of_the_numbers <- all_institutional_facebook_urls_b %>% 
  mutate(id = stringr::str_extract_all(url, '\\(?[0-9,.]+')) %>% 
  pull(id)

find_15_digits <- function(x) {
  x[nchar(x) == 15]
}

length_f <- function(x) {
  length(x) > 0  
}

all_of_the_numbers <- all_of_the_numbers %>% 
  map(possibly(find_15_digits, NA))

all_institutional_facebook_urls_b$id <- map(all_of_the_numbers, function(x) if(identical(x, character(0))) NA_character_ else x)

all_institutional_facebook_urls_b <- all_institutional_facebook_urls_b %>% 
  unnest(id)

x <- all_institutional_facebook_urls_b %>% 
  mutate(url = str_c("https://www.facebook.com/", id)) %>% 
  filter(!is.na(url)) %>% 
  select(-id)

from_15_char_id <- data_with_nces_id_joined_not_matched %>% 
  select(-nces_id) %>% 
  left_join(x) %>% 
  filter(!is.na(nces_id))
```

## joining all back up - all done!

```{r}
all_to_join <- bind_rows(new_joined_data) %>% 
  bind_rows(new_data_without_dashes) %>% 
  bind_rows(from_15_char_id)

all_to_join <- all_to_join %>% 
  distinct(url, post, nces_id, page_name, year)

data_with_nces_id_joined_match_all <- data_with_nces_id_joined_match %>% 
  bind_rows(all_to_join)

str_c("added ", nrow(data_with_nces_id_joined_match_all) - nrow(data_with_nces_id_joined_match), " rows; ",
      (nrow(data_with_nces_id_joined_match_all) - nrow(data_with_nces_id_joined_match)) * 100 / nrow(posts_100k), "% change")
```

# Joining additional NCES data

This data is from the ELSI Table Generator: https://nces.ed.gov/ccd/elsi/tableGenerator.aspx

May want to update to 18-19 data for districts; this came out after I scraped data

```{r}
nces_info_for_districts <- read_csv(here::here("data", "nces-info-for-districts.csv")) %>% 
  select(nces_id, free_and_reduced_lunch_students_public_school_2017_18, state)

nces_info_for_schools <- read_csv("data/ELSI_csv_export_6374077161361122015895.csv") %>% 
  janitor::clean_names() %>% 
  mutate(nces_id = str_extract(school_id_nces_assigned_public_school_latest_available_year, '\\(?[0-9,.]+')) %>% 
  select(nces_id, free_and_reduced_lunch_students_public_school_2018_19, state = state_name_public_school_latest_available_year)

nces_info <- nces_info_for_districts %>% 
  bind_rows(nces_info_for_schools)

nces_info$nces_id %>% 
  nchar() %>% 
  table() # this looks right

data_with_nces_id_joined_match_all$nces_id %>% 
  nchar() %>% 
  table() # so does this

data_with_nces_id_joined_match_all %>% 
  left_join(nces_info, by = "nces_id") %>% 
  filter(is.na(state)) # almost all joined

data_with_nces_id_joined_match_all %>% 
  left_join(nces_info, by = "nces_id") %>% 
  mutate(state = tolower(state)) %>% 
  count(state) %>% 
  knitr::kable() # success!

str_c("Matched ", nrow(data_with_nces_id_joined_match_all)/nrow(posts_100k) * 100, "% of rows with NCES data")
```

# Exploring NCES data

## n accounts

```{r}
nces_dist <- read_csv(here::here("data", "nces-info-for-districts.csv"))

nces_scho <- read_csv("data/ELSI_csv_export_6374077161361122015895.csv") %>% 
  janitor::clean_names() %>% 
  mutate(nces_id = str_extract(school_id_nces_assigned_public_school_latest_available_year, '\\(?[0-9,.]+'))

all_nces_ids <- data_with_nces_id_joined_match_all %>% 
  distinct(nces_id)

n_districts <- nces_dist %>% 
  filter(agency_type_district_2017_18 == "1-Regular local school district that is NOT a component of a supervisory union") %>% 
  semi_join(all_nces_ids) %>% 
  nrow() 

n_districts/8193
```

## checking sample

```{r}
nces_dist$frpl <- as.integer(nces_dist$free_and_reduced_lunch_students_public_school_2017_18) /
  as.integer(nces_dist$total_students_all_grades_excludes_ae_district_2017_18)

mean_fb_frpl <- nces_dist %>% 
  filter(agency_type_district_2017_18 == "1-Regular local school district that is NOT a component of a supervisory union") %>% 
  semi_join(all_nces_ids) %>% 
  summarize(mean_frpl = mean(frpl, na.rm = TRUE))

mean_frpl <- nces_dist %>% 
  filter(agency_type_district_2017_18 == "1-Regular local school district that is NOT a component of a supervisory union") %>% 
  summarize(mean_frpl = mean(frpl, na.rm = TRUE))

mean_fb_frpl
mean_frpl

fb_ss <- nces_dist %>% 
  filter(agency_type_district_2017_18 == "1-Regular local school district that is NOT a component of a supervisory union") %>% 
  semi_join(all_nces_ids) %>% 
  summarize(mean_ss = mean(as.integer(total_students_all_grades_excludes_ae_district_2017_18), na.rm = TRUE))

mean_ss <- nces_dist %>% 
  filter(agency_type_district_2017_18 == "1-Regular local school district that is NOT a component of a supervisory union") %>% 
  summarize(mean_ss = mean(as.integer(total_students_all_grades_excludes_ae_district_2017_18), na.rm = TRUE))

fb_ss
mean_ss
```

## writing data

```{r}
d1 <- data_with_nces_id_joined_match_all %>% 
  left_join(select(nces_dist, -url)) %>% 
  select(url, nces_id, message, state, frpl, n_students = total_students_all_grades_excludes_ae_district_2017_18)

write_csv(d1, "2020-district-data.csv")

d1 <- read_csv("2019-district-data.csv")
d2 <- read_csv("2020-district-data.csv")

bind_rows(d1, d2) %>% write_rds("data/2019-2020-districts.rds")
```

## activity

```{r}
to_join_for_activity <- nces_dist %>% 
  filter(agency_type_district_2017_18 == "1-Regular local school district that is NOT a component of a supervisory union")

activity_d <- data_with_nces_id_joined_match_all %>% 
  count(nces_id) %>% 
  left_join(to_join_for_activity)

activity_d %>% 
  select(n, frpl, total_ss = total_students_all_grades_excludes_ae_district_2017_18) %>% 
  corrr::correlate() %>% 
  knitr::kable()

```

## Posts per account

```{r}
activity_d %>% 
  select(n) %>% 
  skimr::skim()
```

```{r}
activity_d %>% 
  ggplot(aes(x = n)) +
  geom_histogram(bins = 100) +
  ggtitle("posts per account") +
  theme_bw()

activity_d %>% 
  filter(n < 50) %>% 
  ggplot(aes(x = n)) +
  geom_histogram(bins = 100) +
  ggtitle("posts per account for accounts with less than 50 posts") +
  theme_bw() +
  scale_x_continuous(breaks = seq(from = 0, to = 49, by = 2), seq(from = 0, to = 49, by = 2)) +
  xlab(NULL)
```

## by state 

```{r}
fb_state <- nces_dist %>% 
  filter(agency_type_district_2017_18 == "1-Regular local school district that is NOT a component of a supervisory union") %>% 
  semi_join(all_nces_ids) %>% 
  count(state)

all_state <- nces_dist %>% 
  filter(agency_type_district_2017_18 == "1-Regular local school district that is NOT a component of a supervisory union") %>% 
  count(state) %>% 
  rename(total_n = n)

left_join(fb_state, all_state) %>% 
  mutate(prop = n / total_n) %>% 
  rename(fb_n = n) %>% 
  arrange(desc(prop)) %>% 
  knitr::kable()
```

## using all possible NCES IDs (either not available or not joined)

```{r}
all_possible_nces_ids <- all_institutional_facebook_urls %>% 
  distinct(nces_id)

fb_pos_n <- nces_dist %>% 
  filter(agency_type_district_2017_18 == "1-Regular local school district that is NOT a component of a supervisory union") %>% 
  semi_join(all_possible_nces_ids) %>% 
  nrow()

fb_pos_n
fb_pos_n/8193

data_with_nces_id_joined_match_all 
```